| Base de datos de abejas ibéricas \n| Iberian bees database\n

### 2022-10-05 Manuscript/Figures/Figure_3.Rmd
Uploading figs as jpg

# Prepare top 20 records
top_20 <- data %>%
  filter(!Accepted_name == "Apis mellifera") %>%
  group_by(Accepted_name) %>%
  summarise(no_rows = length(Accepted_name)) %>%
  arrange(-no_rows) %>%
  slice(1:20)
# Fix row position
top_20$Accepted_name <- factor(top_20$Accepted_name, levels = top_20$Accepted_name)
# Plot
ggplot(top_20, aes(x = Accepted_name, y = no_rows)) +
  geom_bar(stat = "identity", fill = "#287D8EFF") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 11, face = "italic"), axis.text.y = element_text(size=12),   axis.title=element_text(size=16), axis.line =element_line(size=1)) +
  ylab("Number of records") +
  xlab("Species") +
  scale_y_continuous(expand = c(0, 0))
```

>>>>

# Prepare top 20 records
top_20 <- data %>%
  filter(!Accepted_name == "Apis mellifera") %>%
  group_by(Accepted_name) %>%
  summarise(no_rows = length(Accepted_name)) %>%
  arrange(-no_rows) %>%
  slice(1:20)
# Fix row position
top_20$Accepted_name <- factor(top_20$Accepted_name, levels = top_20$Accepted_name)
# Plot
ggplot(top_20, aes(x = Accepted_name, y = no_rows)) +
  geom_bar(stat = "identity", fill = "#287D8EFF") +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 11, face = "italic"), axis.text.y = element_text(size=12),   axis.title=element_text(size=16), axis.line =element_line(size=1)) +
  ylab("Number of records") +
  xlab("Species") +
  scale_y_continuous(expand = c(0, 0))

ggsave("Figure_3.jpg", dpi=1500)

```

### 2022-10-05 Manuscript/manuscript.Rmd
minor edit
Uploading figs as jpg

In the absence of standardized monitoring efforts, one may think that detailed information on species occurrences in space and time does not exists. However, there are different sources of extremely useful information. First, natural history museums hold historical collections of amateur naturalists and researchers who collected bees [@bartomeus2019]. Collectors are very good at classifying and labeling each collected specimen, which means we can retrieve the information stored along the pinned specimens about when and where those where collected. Second, the Iberian Peninsula has a rich tradition of pollinator and pollination researchers [@archer2014], which have been collecting occurrence bee data with different aims. Finally, many occurrence records are being made available through internet portals which centralize data hosted in institutions all around the world (e.g. GBIF) or collected by citizen scientists (e.g. iNaturalist). The challenge is how to digitalize, access, and curate all these different data sources. Thus, despite the obvious challenges to digitalize, access, and curate all these different data sources, they provide an essential source of information for biodiversity research [@nelson2019; @hedrick2020]

Here we have compiled an Iberian bee database of occurrence records. For this, we have digitalize natural history records from museums, compiled more than 100 datasets contributed by individual researchers, and gathered publicly available information from online repositories. After data cleaning and harmonization, we release the first `r formatC(nrow(data),format="d", big.mark=",")` records from the Iberian Peninsula and the Balearic Islands. **Figure 1** shows the spatial distribution of records, which as expected it is biased to the main urban areas. Overall, the database contains information of `r length(unique(data$Accepted_name))` different species and in **Figure 2** we show the 20 most frequently collected species. The data has been collected between `r min(data$Year, na.rm = T)` and `r max(data$Year, na.rm = T)`, with an increasing trend of the number of records with time, reflecting a renewed interest for documenting bee species (**Fig. 3A**). As expected, most records were recorded in spring with May being the month with highest number of records (**Fig. 3B**). More than `r round(geo/nrow(data),2)*100`% of these records have associated information such as the location and date of capture, and `r round(flw/nrow(data),2)*100`% have extra data such as the plant species where bee specimens were collected. Here, we release the version 1.0 of the Iberian bee database with name "IberianBees". Although this database is a work in progress, we believe that this is a good starting point to better understand and conserve bee biodiversity. It can be accessed at the following permanent link: <https://doi.org/10.5281/zenodo.6354502>.

@@

>>>>

In the absence of standardized monitoring efforts, one may think that detailed information on species occurrences in space and time does not exists. However, there are different sources of extremely useful information. First, natural history museums hold historical collections of amateur naturalists and researchers who collected bees [@bartomeus2019]. Collectors are very good at classifying and labeling each collected specimen, which means we can retrieve the information stored along the pinned specimens about when and where those where collected. Second, the Iberian Peninsula has a rich tradition of pollinator and pollination researchers [@archer2014], which have been collecting occurrence bee data with different aims. Finally, many occurrence records are being made available through internet portals which centralize data hosted in institutions all around the world (e.g. Global Biodiversity Information Facility, GBIF) or collected by citizen scientists (e.g. iNaturalist). The challenge is how to digitalize, access, and curate all these different data sources. Thus, despite the obvious challenges to digitalize, access, and curate all these different data sources, they provide an essential source of information for biodiversity research [@nelson2019; @hedrick2020]

Here we have compiled an Iberian bee database of occurrence records. For this, we have digitalize natural history records from museums, compiled more than 100 datasets contributed by individual researchers, and gathered publicly available information from online repositories (see **Fig. 1** for workflow). After data cleaning and harmonization, we release the first `r formatC(nrow(data),format="d", big.mark=",")` records from the Iberian Peninsula and the Balearic Islands. **Figure 2** shows the spatial distribution of records, which as expected it is biased to the main urban areas. Overall, the database contains information of `r length(unique(data$Accepted_name))` different species and in **Figure 3** we show the 20 most frequently collected species. The data has been collected between `r min(data$Year, na.rm = T)` and `r max(data$Year, na.rm = T)`, with an increasing trend of the number of records with time, reflecting a renewed interest for documenting bee species (**Fig. 4A**). As expected, most records were recorded in spring with May being the month with highest number of records (**Fig. 4B**). More than `r round(geo/nrow(data),2)*100`% of these records have associated information such as the location and date of capture, and `r round(flw/nrow(data),2)*100`% have extra data such as the plant species where bee specimens were collected. Here, we release the version 1.0 of the Iberian bee database with name "IberianBees". Although this database is a work in progress, we believe that this is a good starting point to better understand and conserve bee biodiversity. It can be accessed at the following permanent link: <https://doi.org/10.5281/zenodo.6354502>.

@@

```{r, out.width="100%", include=TRUE, fig.align="center", echo=FALSE}
knitr::include_graphics("Figures/Figure_1.pdf")
```
